---
title: "Prototyping_I"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prototyping_I}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(MinervaTime)
```


## Jack in the box

The toy jack-in-the-box is a box with a handle. Turning the handle plays music for an unpredictable amount of time, followed by jack popping out of the box.

The situation is analogous to simple acquisition Pavlovian conditioning involving repeated pairing of a cue (e.g., tone or light) and outcome (e.g., food reward).

MINERVA-AL was applied as an instance-theory approach to modeling associative learning phenomena like classical conditioning. However, MINERVA-AL is a trial-level model and does not address any role for temporal dynamics or expectation. For example, MINERVA-AL could be trained on several rounds of playing jack-in-the-box, which would bye the equivalent of many CS+ trials. As a result, MINERVA-AL would learn that given the box, jack will pop out. However, MINERVA-AL would have no within-trial expectation about when Jack will pop out.

The purpose of this modelling effort is to develop a MINERVA account of intra-trial temporal expectations, and inter-trial general expectations.

## Environment vectors

MINERVA-AL represents cues and outcomes in a concatenated feature vector. For example, the first 10 numbers represent the presence or absences of an A feature, the second set of 10 number represent the presence or absence of a B feature, and so on. If A is followed by a reward, then the environmental feature vector would have +1s in the A field, and +1s in the outcome field, and 0s in the other fields. Notably, there is no representation of time between the presentation of the cue and outcome.

We will add a representation of time directly as part of the feature vector. In this way, temporal context features can be represented in memory, and can influence retrieval and echo construction.

We could use a general syntax for generating the environment vectors describing the temporal location of the appearance, duration, and disappearance of cues and outcomes. Something like a MIDI syntax that can be transformed to MINERVA vectors.

### Quick start example

```{r}

veclength <- 10 

cue <- list(features = c(rep(1,veclength)),
            onset = 1,
            duration = 100,
            offset = 100
)
outcome <- list(features = c(rep(1,veclength)),
            onset = 50,
            duration = 50,
            offset = 100
)
context <- list(features = c(rep(1,veclength)),
            onset = 1,
            duration = 200,
            offset = 200
)

timesteps <- 1:200

# generate matrix of temporal vectors
# timepoints closer together in time are more similar
temporal_matrix <- matrix(1, 
                          nrow = length(timesteps),
                          ncol = length(timesteps))

lower <- -2
upper <- 2
delta <- row(temporal_matrix) - col(temporal_matrix)
temporal_matrix[delta < lower | delta > upper] <- 0

# cues
cue_matrix <- matrix(0,
                     nrow=length(timesteps),
                     ncol=length(cue$features))

cue_matrix[cue$onset:cue$offset,] <- cue$features

# outcome
outcome_matrix <- matrix(0,
                     nrow=length(timesteps),
                     ncol=length(outcome$features))

outcome_matrix[outcome$onset:outcome$offset,] <- outcome$features

# context
context_matrix <- matrix(0,
                     nrow=length(timesteps),
                     ncol=length(context$features))

context_matrix[context$onset:context$offset,] <- context$features

# Whole Environment Matrix

environment_matrix <- cbind(cue_matrix,
                            outcome_matrix,
                            context_matrix,
                            temporal_matrix)

```

### Run the model

```{r}
library(RsemanticLibrarian)

# initialize memory with 5 noise vectors
noise <- matrix(runif(5*dim(environment_matrix)[2],-1,1)*.05,
                 nrow = 5,
                 ncol = dim(environment_matrix)[2])

memory <- matrix(0,
                 nrow=dim(environment_matrix)[1],
                 ncol=dim(environment_matrix)[2])
memory <- rbind(noise,memory)

tau <- 3

save_echo <- matrix(0,
                    nrow=dim(environment_matrix)[1],
                    ncol=dim(environment_matrix)[2])

# intra-trial encoding and retrieval
for (i in 1:dim(environment_matrix)[1]){
  
  activations <- c(cosine_x_to_m(environment_matrix[i,],
                                 memory[1:(4+i),])^tau)
  echo <- colSums(memory[1:(4+i),]*activations)
  save_echo[i,] <- echo
  memory[5+i,] <- environment_matrix[i,]

}

```

### evaluate

```{r}

cue_expectation <- rowMeans(save_echo[,1:10])
plot(cue_expectation)

outcome_expectation <- rowMeans(save_echo[,11:20])
plot(outcome_expectation)

```

## Event Representation

In the above example, a cue and outcome event were presented over slices of 200 time points. This section develops a more general approach to describing event representation. 

A trial will refer to a series of time points where events appear and disappear for particular durations, like musical notes placed within a bar.

A timeline will refer to a series of trials. 

Notepad:

- make_trial: a function to name each event and associated onset, duration, offset parameters
- make_timeline: a function to write a series of trials to a dataframe...Not clear whether a function is useful here. Easier to code specific timelines, possibly useful to create specific timeline functions that are limited to particular cases. E.g., A appears followed by B, where B's onset is defined by a distribution.
- vectorize_timeline: a function to create environmental vectors for the trials andevents in a timeline


```{r, eval = FALSE}
library(dplyr)

# matrix of event feature vectors
event_vectors <- matrix(1, ncol = 10, nrow = 10)
row.names(event_vectors) <- LETTERS[1:10]

# make an event frame, define the events for a single timeslice
make_trial <- function(trial_num,
                       event_names,
                       onsets,
                       durations) {

  trial <- data.frame(
    trial_num = trial_num,
    event = event_names,
    onset = onsets,
    duration = durations,
    offset = onsets+durations
  )

  return(trial)
}

# test function
#make_trial(1,event_names=c("A","B"),c(1,100),durations=50)

# create a timeline of events using a data.frame
timeline <- data.frame()

for(i in 1:5){
  new_trial <- make_trial(i,event_names=c("A","B"),c(1,100),durations=50)
  timeline <- rbind(timeline,new_trial)
}

# a simple timeline function for generating AB trials
make_AB_timeline <- function(num_trials =5,
                                 events = c('A','B'),
                                 A_onsets = 1,
                                 A_durations = 50,
                                 B_onsets = 50,
                                 B_durations = 50
                                 ){
  timeline <- data.frame()
  
  for(i in 1:num_trials){
    
    A_on <- A_onsets[1]
    if(length(A_onsets) > 1) A_on <- A_onsets[i]
    B_on <- B_onsets[1]
    if(length(B_onsets) > 1) B_on <- B_onsets[i]

    new_trial <- make_trial(i,
                            event_names=events,
                            onsets=c(A_on,B_on),
                            durations=c(A_durations,B_durations))
    
    timeline <- rbind(timeline,new_trial)
  }
  
  return(timeline)
  
}

make_AB_timeline(
  num_trials = 5,
  events = c('A', 'B'),
  A_onsets = 1,
  A_durations = 50,
  B_onsets = round(rnorm(5,100,10)),
  B_durations = 50
)

## Another simple timeline

# a simple timeline function for generating AB trials
make_event_timeline <- function(num_trials = 5,
                                 events = c('A','B','C'),
                                 onsets = list(1,round(rnorm(5,50,10)),1),
                                 durations = list(25,25,199)
                                 ){
  timeline <- data.frame()
  
  for(i in 1:num_trials){
    
    # could functionalize this later
    process_onsets <- onsets
    for(o in 1:length(process_onsets)){
      if(length(process_onsets[[o]]) == 1) {
        process_onsets[[o]] <- rep(process_onsets[[o]],num_trials)
      }
    }
    process_onsets <- matrix(unlist(process_onsets),nrow=num_trials, byrow=FALSE)
    
    process_durations <- durations
    for(d in 1:length(process_durations)){
      if(length(process_durations[[d]]) == 1) {
        process_durations[[d]] <- rep(process_durations[[d]],num_trials)
      }
    }
    process_durations <- matrix(unlist(process_durations),nrow=num_trials, byrow=FALSE)

    new_trial <- make_trial(i,
                            event_names=events,
                            onsets=process_onsets[i,],
                            durations=process_durations[i,])
    
    timeline <- rbind(timeline,new_trial)
  }
  
  return(timeline)
  
}

make_event_timeline(
  num_trials = 5,
  events = c('A', 'B', 'C'),
  onsets = list(1,rnorm(5,50,10),1),
  durations = list(25,25,199)
)

# turn a timeline into a matrix of environmental vectors

# create a matrix of temporal vectors

make_temporal_vectors <- function(num_timesteps = 200, overlap = 2){
  
temporal_matrix <- matrix(1, 
                          nrow = num_timesteps,
                          ncol = num_timesteps)

lower <- -1*overlap
upper <- overlap
delta <- row(temporal_matrix) - col(temporal_matrix)
temporal_matrix[delta < lower | delta > upper] <- 0

return(temporal_matrix)
  
}

temporal_vectors <- make_temporal_vectors()

# make event_vectors

make_event_vectors <- function(event_names = LETTERS[1:10],
                               vector_length = 10) {
  event_vectors <- matrix(1,
                          ncol = vector_length,
                          nrow = length(event_names))
  
  row.names(event_vectors) <- event_names
  
  return(event_vectors)
  
}

event_vectors <- make_event_vectors(LETTERS[1:10],10)

# Vectorize the timeline

timeline_to_vector <- function(timeline,
                               event_vectors,
                               temporal_vectors,
                               timesteps = 200) {
  timeline_list <- list()
  
  # iterate through each trial
  for(t in 1:max(timeline$trial_num)){
    
    # get events for current trial
    current_trial <- timeline[timeline$trial_num == t,]
    
    # create event vector matrix
    all_events_matrix <- matrix(nrow = timesteps)
    for (e in 1:length(current_trial$event)){
      e_matrix <- matrix(0, 
                         nrow = timesteps, 
                         ncol = length(event_vectors[current_trial[e,'event'],]))
      
      e_matrix[(current_trial[e,'onset']:current_trial[e,'offset']),] <- event_vectors[current_trial[e,'event'],]
      
      #append matrix for each event to whole trial matrix
     all_events_matrix <-  cbind(all_events_matrix,e_matrix)
    }
    all_events_matrix <- all_events_matrix[,-1] # delete column of NAs
    
    # append temporal vectors
    all_events_matrix <- cbind(all_events_matrix,temporal_vectors)
    
    # save to list
    timeline_list[[t]] <- all_events_matrix
  }
  
  return(timeline_list)
}

timeline_vectors <- timeline_to_vector(timeline,
                                       event_vectors,
                                       temporal_vectors,200)

```


## Create trials

```{r}

#### BUILD TRAINING SCENARIO
# e.g., a matrix of environment vectors
# representing events unfolding over time within a trial
# with the capability to train multiple trials


trial_timesteps <- 200 # define number of time steps per trial

# create a timeline of trials, listing temporal event properties per trial
timeline <- make_event_timeline(
 num_trials = 10,
 events = c('A', 'B', 'C'),
 onsets = list(1,rnorm(10,50,10),1),
 durations = list(25,25,199)
 )

# convert timeline to environment vectors
temporal_vectors <- make_temporal_vectors(trial_timesteps, overlap = 2)
event_vectors <- make_event_vectors(LETTERS[1:10],10)

timeline_vectors <- timeline_to_vector(timeline,
                                       event_vectors,
                                       temporal_vectors,200)


```

## Run model

```{r}
## define basic model parameters

library(RsemanticLibrarian) # for cosine_x_to_m

# initialize memory with 5 noise vectors
noise <- matrix(runif(5*dim(environment_matrix)[2],-1,1)*.05,
                 nrow = 5,
                 ncol = dim(environment_matrix)[2])

#memory <- matrix(0,
#                 nrow=dim(environment_matrix)[1],
#                 ncol=dim(environment_matrix)[2])
memory <- noise

tau <- 3

## Run the model for each trial

model_results <- list()

for(t in 1:length(timeline_vectors)) {
  environment_matrix <- timeline_vectors[[t]]
  
  save_echo <- matrix(0,
                      nrow=dim(environment_matrix)[1],
                      ncol=dim(environment_matrix)[2])
  
  # intra-trial encoding and retrieval
  for (i in 1:dim(environment_matrix)[1]){
    
    activations <- c(cosine_x_to_m(environment_matrix[i,],
                                   memory)^tau)
    echo <- colSums(memory*activations)
    save_echo[i,] <- echo
    memory <- rbind(memory,environment_matrix[i,])
  
  }
  
  A_expectation <- rowMeans(save_echo[,1:10])
  #plot(A_expectation)
  
  B_expectation <- rowMeans(save_echo[,11:20])
  #plot(B_expectation)
  
  model_results[[t]] <- list(A_expectation = A_expectation,
                             B_expectation = B_expectation,
                             save_echo = save_echo)

}

```

```{r, eval= FALSE}

timeline_to_vector2 <- function(timeline,
                               event_vectors,
                               temporal_vectors,
                               timesteps = 200) {
  timeline_list <- list()

  # iterate through each trial
  for(t in 1:max(timeline$trial_num)){

    # get events for current trial
    current_trial <- timeline[timeline$trial_num == t,]

    # create event vector matrix
    all_events_matrix <- matrix(0,
                                nrow = timesteps,
                                ncol = dim(event_vectors)[1]*dim(event_vectors)[2])
    
    # loop through each event name in trial
    for (e in 1:length(current_trial$event)){
      
      # create blank matrix for event
      e_matrix <- matrix(0,
                         nrow = timesteps,
                         ncol = length(event_vectors[current_trial[e,'event'],]))
      
      # add event vector for specified durations in matrix
      e_matrix[(current_trial[e,'onset']:current_trial[e,'offset']),] <- event_vectors[current_trial[e,'event'],]

      #append matrix for each event to whole trial matrix
      #all_events_matrix <-  cbind(all_events_matrix,e_matrix)
      
      #insert event matrix to it's field position
      e_name <- current_trial[e,'event']
      e_row <- which(row.names(event_vectors) %in% e_name)
      e_length <- length(event_vectors[e_row,])
      first_ind <- ((e_row-1)*e_length)+1
      last_ind <- first_ind + (e_length-1)
      all_events_matrix[ ,first_ind:last_ind] <- e_matrix

    }
    
   # all_events_matrix <- all_events_matrix[,-1] # delete column of NAs

    # append temporal vectors
    all_events_matrix <- cbind(all_events_matrix,temporal_vectors)

    # save to list
    timeline_list[[t]] <- all_events_matrix
  }

  return(timeline_list)
}


trial_timesteps <- 200 # define number of time steps per trial

# create a timeline of trials, listing temporal event properties per trial
timeline <- make_event_timeline(
 num_trials = 10,
 events = c('A', 'B', 'C'),
 onsets = list(1,round(rnorm(10,50,10)),1),
 durations = list(25,25,199)
 )

# convert timeline to environment vectors
temporal_vectors <- make_temporal_vectors(trial_timesteps, overlap = 2)
event_vectors <- make_event_vectors(LETTERS[1:10],10)

timeline_vectors <- timeline_to_vector2(timeline,
                                       event_vectors,
                                       temporal_vectors,200)
View(timeline_vectors[[2]])


```



