---
title: "Working Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{working_example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MinervaTime)
```

See the reference section for notes on individual functions.

This working example contains 10 trials. On each trial an A event is presented for 25 time steps, followed by a B event for 25 time steps. The onset of the B event is randomly determined by a normal distribution, centered on the 50th time step, with a standard deviation of 10 timesteps.

## Create trials

```{r}

#### BUILD TRAINING SCENARIO
# e.g., a matrix of environment vectors
# representing events unfolding over time within a trial
# with the capability to train multiple trials


trial_timesteps <- 200 # define number of time steps per trial

# create a timeline of trials, listing temporal event properties per trial
timeline <- make_event_timeline(
 num_trials = 10,
 events = c('A', 'B', 'C'),
 onsets = list(1,rnorm(10,50,10),1),
 durations = list(25,25,199)
 )

# convert timeline to environment vectors
temporal_vectors <- make_temporal_vectors(trial_timesteps, overlap = 2)
event_vectors <- make_event_vectors(LETTERS[1:10],10)

timeline_vectors <- timeline_to_vector(timeline,
                                       event_vectors,
                                       temporal_vectors,200)


```

## Run model

```{r}
## define basic model parameters

library(RsemanticLibrarian) # for cosine_x_to_m

# initialize memory with 5 noise vectors
environment_matrix <- timeline_vectors[[1]]
noise <- matrix(runif(5*dim(environment_matrix)[2],-1,1)*.05,
                 nrow = 5,
                 ncol = dim(environment_matrix)[2])

#memory <- matrix(0,
#                 nrow=dim(environment_matrix)[1],
#                 ncol=dim(environment_matrix)[2])
memory <- noise

tau <- 3

## Run the model for each trial

model_results <- list()

for(t in 1:length(timeline_vectors)) {
  environment_matrix <- timeline_vectors[[t]]
  
  save_echo <- matrix(0,
                      nrow=dim(environment_matrix)[1],
                      ncol=dim(environment_matrix)[2])
  
  # intra-trial encoding and retrieval
  for (i in 1:dim(environment_matrix)[1]){
    
    activations <- c(cosine_x_to_m(environment_matrix[i,],
                                   memory)^tau)
    echo <- colSums(memory*activations)
    save_echo[i,] <- echo
    memory <- rbind(memory,environment_matrix[i,])
  
  }
  
  A_expectation <- rowMeans(save_echo[,1:10])
  #plot(A_expectation)
  
  B_expectation <- rowMeans(save_echo[,11:20])
  #plot(B_expectation)
  
  model_results[[t]] <- list(A_expectation = A_expectation,
                             B_expectation = B_expectation,
                             save_echo = save_echo)

}

```

## plot model results

Shows the model's expectation for the B event across a trial window, after the 1st, 5th, and 10th trial.

```{r}
plot(model_results[[1]]$B_expectation)
plot(model_results[[5]]$B_expectation)
plot(model_results[[10]]$B_expectation)
```

